To support formalization in type theory of research level mathematics in universal algebra and related fields, we present the Agda Universal Algebra Library (\agdaualib), a software library containing formal statements and proofs of the core definitions and results of universal algebra. 
The \ualib is written in \agda~\cite{Norell:2009}, a programming language and proof assistant based on Martin-L\"of Type Theory that not only supports dependent and inductive types, as well as proof tactics for proving things about the objects that inhabit these types.

There have been a number of efforts to formalize parts of universal algebra in type theory prior to ours, most notably
\begin{itemize}
  \item Capretta~\cite{Capretta:1999} (1999) formalized the basics of universal algebra in the Calculus of Inductive Constructions using the Coq proof assistant;
    \item Spitters and van der Weegen~\cite{Spitters:2011} (2011) formalized the basics of universal algebra and some classical algebraic structures, also in the Calculus of Inductive Constructions using the Coq proof assistant, promoting the use of type classes as a preferable alternative to setoids;
 \item Gunther, et al~\cite{Gunther:2018} (2018) developed what seems to be (prior to the \ualib) the most extensive library of formal universal algebra to date; in particular, this work includes a formalization of some basic equational logic; the project (like the \ualib) uses Martin-L\"of Type Theory and the Agda proof assistant.
\end{itemize}
Some other projects aimed at formalizing mathematics generally, and algebra in particular, have developed into very extensive libraries that include definitions, theorems, and proofs about algebraic structures, such as groups, rings, modules, etc.  However, the goals of these efforts seem to be the formalization of special classical algebraic structures, as opposed to the general theory of (universal) algebras. Moreover, the part of universal algebra and equational logic formalized in the \ualib extends beyond the scope of prior efforts and. In particular, the library now includes a proof of Birkhoff's variety theorem.  Most other proofs of this theorem that we know of are informal and nonconstructive.\footnote{After the completion of this work, the author learned about a constructive version of Birkhoff's theorem that was proved by Carlstr\"om in~\cite{Carlstrom:2008}.  The latter is presented in the standard, informal style of mathematical writing in the current literature, and as far as we know it was never implemented formally and type-checked with a proof assistant. Nonetheless, a comparison of the version of the theorem presented in~\cite{Carlstrom:2008} to the Agda proof we give here would be interesting.}
 %% We remark briefly on this in~\S\ref{sec:conclusions-and-future-work}.}

%% \subsection{Vision and Goals}\label{vision-and-goals}
The seminal idea for the \agdaualib project was the observation that, on the one hand, a number of fundamental constructions in universal algebra can be defined recursively, and theorems about them proved by structural induction, while, on the other hand, inductive and dependent types make possible very precise formal representations of recursively defined objects, which often admit elegant constructive proofs of properties of such objects.  An important feature of such proofs in type theory is that they are total functional programs and, as such, they are computable, composable, and machine-verifiable.
%% These observations suggested that there would be much to gain from implementing universal algebra in a language, such as Martin-L\"of type theory, that features dependent and inductive types.

Finally, our own research experience has taught us that a proof assistant and programming language (like Agda), when equipped with specialized libraries and domain-specific tactics to automate the proof idioms of our field, can be an extremely powerful and effective asset. As such we believe that proof assistants and their supporting libraries will eventually become indispensable tools in the working mathematician's toolkit.



\subsection{Contributions and organization}
\label{sec:contributions}

In this paper we limit ourselves to the presentation of the core foundational modules of the \ualib so that we have space to discuss some of the more interesting type theoretic and foundational issues that arose when developing the library and attempting to represent advanced mathematical notions in type theory and formalize them in Agda. We are in the process of writing two follow-up papers covering more advanced aspects of the library. The contents of these sequels are mentioned in the conclusion of the paper.

This present paper is organized into three parts as follows. The first part is \S\ref{sec:agda-prelude} which introduces the basic concepts of type theory with special emphasis on the way such concepts are formalized in \agda. Specifically, \S\ref{preliminaries} introduces \emph{Sigma types} and Agda's hierarchy of \emph{universes}. The important topics of \emph{equality} and \emph{function extensionality} are discussed in \S\ref{equality} and \S\ref{function-extensionality}; \S\ref{sec:inverse-image-invers} covers inverses and inverse images of functions. In \S\ref{sec:lifts-altern-univ} we describe a technical problem that one frequently encounters when working in a \emph{noncummulative universe hierarchy} and offer some tools for resolving the type-checking errors that arise from this.

The second part (\S\ref{sec:relat-quot-types}) covers relations and quotient types. Specifically, \S\ref{sec:discrete-relations} defines types that represent \emph{unary} and \emph{binary relations} as well as \emph{function kernels}. These ``discrete relation types,'' are all very standard.  In \S\ref{sec:continuous-relations} we introduce the (less standard) types that we use to represent \emph{general} and \emph{dependent relations}. We call these ``continuous relations'' because they can have arbitrary arity (general relations) and they can be defined over arbitrary families of types (dependent relations).
%
Section~\ref{sec:equiv-relat-quot} covers standard types for equivalence relations and quotients, and \S\ref{sec:trunc-sets-prop} discusses a family of concepts that are vital to the mechanization of mathematics using type theory; these are the closely related concepts of \emph{truncation}, \emph{sets}, \emph{propositions}, and \emph{proposition extensionality}.

The third part of the paper is \S\ref{sec:algebra-types} which covers the basic domain-specific types offered by the \ualib. It is here that we finally get to see some types representing algebraic structures.  Specifically, we describe types for \emph{operations} and \emph{signatures} (\S\ref{sec:oper-sign}), \emph{general algebras} (\S\ref{sec:algebras}), and \emph{product algebras}, including types for representing \emph{products over arbitrary classes of algebraic structures} (\S\ref{sec:product-algebras}). Finally, we define types for congruence relations and quotient algebras in \SS\ref{congruences}--\ref{quotient-algebras}.

To conclude this introduction, we offer the following links to official sources of information about the \agdaualib.
\begin{itemize}
  \item \href{https://ualib.gitlab.io}{\texttt{ualib.org}} (the web site) includes every line of code in the library, rendered as html and accompanied by documentation, and
  \item \href{https://gitlab.com/ualib/ualib.gitlab.io}{\texttt{gitlab.com/ualib/ualib.gitlab.io}} (the source code) freely available and licensed under the \href{https://creativecommons.org/licenses/by-sa/4.0/}{Creative Commons Attribution-ShareAlike 4.0 International License}.
\end{itemize}
These sources include complete proofs of every theorem we mention here, and much more, including the Agda modules that will be covered in the second and third installments in this series of papers on the \ualib.

Finally, we believe readers of this paper will get much more out of it if they download the software from \url{https://gitlab.com/ualib/ualib.gitlab.io/}, install the library, and try it out for themselves.
